Comparing binsegRcpp with other implementations of binary segmentation

### 2022-04-07 article.Rnw
container tab
meanvar_norm timings redo

%% - \Abstract{} almost as usual
\Abstract{ Binary segmentation is a classic algorithm for detecting
 changepoints in sequential data. In theory it should be extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst case, and $O(N \log K)$ in the best case. In practice existing implementations such as ruptures (\proglang{Python} module) and changepoint (\proglang{R} package) are much slower, and in fact sometimes do not return a correct result. We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient C++ implementation of binary segmentation, and always returns correct results. We discuss some advanced C++ coding techniques which were used to avoid repetition, for efficiency, and for readability.  We additionally include detailed theoretical and empirical comparisons with other implementations of binary segmentation in \proglang{R} packages \pkg{fpop} and \pkg{wbs}. }

@@

>>>>

%% - \Abstract{} almost as usual
\Abstract{ Binary segmentation is a classic algorithm for detecting
 changepoints in sequential data. In theory it should be extremely fast for $N$ data and $K$ segments, $O(N K)$ in the worst case, and $O(N \log K)$ in the best case. In practice existing implementations such as ruptures (\proglang{Python} module) and changepoint (\proglang{R} package) are much slower, and in fact sometimes do not return a correct result. We present the \proglang{R} package \pkg{binsegRcpp}, which provides an efficient C++ implementation of binary segmentation, and always returns correct results. We discuss some advanced C++ coding techniques which were used to avoid repetition, for efficiency, and for readability.  We additionally include detailed theoretical and empirical comparisons with other implementations of binary segmentation in \proglang{R} packages \pkg{fpop}, \pkg{blockcpd} and \pkg{wbs}. }

@@

\begin{table}
  \centering
  \begin{tabular}{c|cc|cc|cc}
    case/splits: & \multicolumn{2}{c}{one } & \multicolumn{2}{c}{best/equal} & \multicolumn{2}{c}{worst/unequal } \\
    operation: & insert &
                             %remove &
                                      argmin & insert & argmin & insert & argmin\\
    \hline
    list & $O(1)$ &
    % $O(1)$ &
               $O(n)$ &
    $O(S\log S)$ & $O(S)$ & $O(S)$ & $O(S)$\\
    heap/multiset & $O(\log n)$ &
    % $O(\log n)$ &
                    $O(1)$ &
                             $O(S)$ & $O(S^2)$ & $O(S)$ & $O(S)$\\
    \hline
    Find new split & \multicolumn{2}{c}{} & \multicolumn{2}{c}{$O(N\log S)$} & \multicolumn{2}{c}{$O(NS)$}
  \end{tabular}
  \caption{container properties}
  \label{tab:containers}
\end{table}
